---
title: "Alpha/Beta Diversities"
author: 
  - Huang Lin$^1$
  - $^1$NICHD, 6710B Rockledge Drive, Bethesda, MD 20892
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document: 
    toc: true
    theme: united
---

```{r setup, include=FALSE}
# rm(list = ls())
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, 
                      message = FALSE, comment = NA,
                      fig.width = 6.25, fig.height = 5)

library(readr)
library(openxlsx)
library(tidyverse) 
library(microbiome)
library(vegan)
library(compositions)
library(magrittr)
library(qwraps2)
library(ggpubr)
```

```{r data}
# OTU table
otu_table = read_tsv("../data/table-l7.txt", skip = 1)
otu_table = otu_table %>% 
  dplyr::select(-taxonomy)
otu_id = otu_table$`#OTU ID`
otu_table = data.frame(otu_table[, -1], check.names = FALSE)
rownames(otu_table) = otu_id

# Taxonomy table
tax = data.frame(Taxon = otu_id)
rownames(tax) = otu_id
tax = tax %>%
  separate(col = Taxon, 
           into = c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"),
           sep = ";")
for (i in 1:ncol(tax)) {
  tax[, i] = sapply(tax[, i], function(x) str_split(x, "__")[[1]][2])
}
tax = as.matrix(tax)
tax[tax == ""] = NA

# Meta data
meta_data = read.xlsx("../data/metadata.xlsx", sheet = 1, colNames = TRUE)
meta_data = meta_data %>%
  transmute(sampleid = Sample.ID, gender = gender, antibiotics,
            age = as.numeric(AGE), 
            age_group = ifelse(adult == 0, "Adolescent", "Adult"),
            bmi = as.numeric(cd2o) / (as.numeric(cd1o) / 100)^2,
            edu = as.character(educationo),
            time_urine = as.numeric(time_urine), batch = New2019, smoke = smoke,
            DCP24_SG = as.numeric(DCP24_SG),
            BP3_SG = as.numeric(BP3_SG), BPA_SG = as.numeric(BPA_SG),
            BPF_SG = as.numeric(BPF_SG), BPS_SG = as.numeric(BPS_SG),
            EPB_SG = as.numeric(EPB_SG), PBP_SG = as.numeric(PBP_SG),
            MPB_SG = as.numeric(MPB_SG), parabensum_SG = as.numeric(parabensum_mol),
            DCP25 = ifelse(belowDCP25 == 0, "Above LOD", "Below LOD"),
            B_PB = ifelse(belowB_PB == 0, "Above LOD", "Below LOD"), 
            TCC = ifelse(belowTCC == 0, "Above LOD", "Below LOD"),
            TCS = ifelse(belowTCS == 0, "Above LOD", "Below LOD"))
rownames(meta_data) = meta_data$sampleid
meta_data$gender = recode(meta_data$gender, 'female' = "Female", `male` = "Male")
meta_data$DCP25 = factor(meta_data$DCP25, levels = c("Below LOD", "Above LOD"))
meta_data$B_PB = factor(meta_data$B_PB, levels = c("Below LOD", "Above LOD"))
meta_data$TCC = factor(meta_data$TCC, levels = c("Below LOD", "Above LOD"))
meta_data$TCS = factor(meta_data$TCS, levels = c("Below LOD", "Above LOD"))

# Combine them into a phyloseq object
OTU = otu_table(otu_table, taxa_are_rows = TRUE)
META = sample_data(meta_data)
TAX = tax_table(tax)
otu_data = phyloseq(OTU, TAX, META)
# Exclude all participants who have used antibiotics in the last four weeks
otu_data = subset_samples(otu_data, antibiotics != "last4weeks")

# Split by age
otu_data1 = subset_samples(otu_data, age_group == "Adult")
otu_data2 = subset_samples(otu_data, age_group == "Adolescent")

# Adults
meta_data1 = meta(otu_data1)
meta_data1 = meta_data1 %>%
  mutate(DCP24 = cut(DCP24_SG, quantile(DCP24_SG, seq(0, 1, 0.25), na.rm = TRUE), labels = 0:3),
         BP3 = cut(BP3_SG, quantile(BP3_SG, seq(0, 1, 0.25), na.rm = TRUE), labels = 0:3),
         BPA = cut(BPA_SG, quantile(BPA_SG, seq(0, 1, 0.25), na.rm = TRUE), labels = 0:3),
         BPF = cut(BPF_SG, quantile(BPF_SG, seq(0, 1, 0.25), na.rm = TRUE), labels = 0:3),
         BPS = cut(BPS_SG, quantile(BPS_SG, seq(0, 1, 0.25), na.rm = TRUE), labels = 0:3),
         EPB = cut(EPB_SG, quantile(EPB_SG, seq(0, 1, 0.25), na.rm = TRUE), labels = 0:3),
         PBP = cut(PBP_SG, quantile(PBP_SG, seq(0, 1, 0.25), na.rm = TRUE), labels = 0:3),
         MPB = cut(MPB_SG, quantile(MPB_SG, seq(0, 1, 0.25), na.rm = TRUE), labels = 0:3),
         parabensum = cut(parabensum_SG, quantile(parabensum_SG, seq(0, 1, 0.25), 
                                                  na.rm = TRUE), labels = 0:3))
rownames(meta_data1) = meta_data1$sampleid
meta_data1 = meta_data1 %>%
  mutate_if(is.numeric, function(x) round(x, 2))

otu_data1 = phyloseq(otu_table(otu_data1), 
                     tax_table(otu_data1), 
                     sample_data(meta_data1))

# Adolescents
meta_data2 = meta(otu_data2)
meta_data2 = meta_data2 %>%
  mutate(DCP24 = cut(DCP24_SG, quantile(DCP24_SG, seq(0, 1, 0.25), na.rm = TRUE), labels = 0:3),
         BP3 = cut(BP3_SG, quantile(BP3_SG, seq(0, 1, 0.25), na.rm = TRUE), labels = 0:3),
         BPA = cut(BPA_SG, quantile(BPA_SG, seq(0, 1, 0.25), na.rm = TRUE), labels = 0:3),
         BPF = cut(BPF_SG, quantile(BPF_SG, seq(0, 1, 0.25), na.rm = TRUE), labels = 0:3),
         BPS = cut(BPS_SG, quantile(BPS_SG, seq(0, 1, 0.25), na.rm = TRUE), labels = 0:3),
         EPB = cut(EPB_SG, quantile(EPB_SG, seq(0, 1, 0.25), na.rm = TRUE), labels = 0:3),
         PBP = cut(PBP_SG, quantile(PBP_SG, seq(0, 1, 0.25), na.rm = TRUE), labels = 0:3),
         MPB = cut(MPB_SG, quantile(MPB_SG, seq(0, 1, 0.25), na.rm = TRUE), labels = 0:3),
         parabensum = cut(parabensum_SG, quantile(parabensum_SG, seq(0, 1, 0.25), 
                                                  na.rm = TRUE), labels = 0:3))
rownames(meta_data2) = meta_data2$sampleid
meta_data2 = meta_data2 %>%
  mutate_if(is.numeric, function(x) round(x, 2))

otu_data2 = phyloseq(otu_table(otu_data2), 
                     tax_table(otu_data2), 
                     sample_data(meta_data2))
```

# 2. Alpha diversity: Shannon’s diversity index

1. For the adult population.

2. P-value is obtained by Wilcoxon Rank Sum Test for alpha diversities.

```{r}
# Aggregate taxa
genus_data1 = aggregate_taxa(otu_data1, "Genus")
d_alpha = alpha(genus_data1, index = "diversity_shannon")

df_alpha = data.frame(meta(genus_data1), 
                      shannon = d_alpha$diversity_shannon)
```

## 2.1 Main figures

```{r, fig.height=8, fig.width=8}
my_comparisons = list(c("0", "1"), c("0", "2"), c("0", "3"),
                      c("1", "2"), c("1", "3"), c("2", "3"))
p_mpb = df_alpha %>%
  filter(!is.na(MPB)) %>%
  ggplot(aes(x = MPB, y = shannon)) + 
  geom_boxplot(width = 0.5, aes(fill = MPB), alpha = 0.7) +
  geom_jitter(shape = 16, position = position_jitter(0.2), size = 0.7) + 
  labs(x = NULL, y = "Shannon diversity index", title = "MPB") + 
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "bottom") +
  stat_compare_means(comparisons = my_comparisons) +
      scale_fill_brewer(name = NULL, palette = "Dark2",
                  label = c("1Q", "2Q", "3Q", "4Q"))

p_ppb = df_alpha %>%
  filter(!is.na(PBP)) %>%
  ggplot(aes(x = PBP, y = shannon)) + 
  geom_boxplot(width = 0.5, aes(fill = PBP), alpha = 0.7) +
  geom_jitter(shape = 16, position = position_jitter(0.2), size = 0.7) + 
  labs(x = NULL, y = "Shannon diversity index", title = "PPB") + 
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "bottom") +
  stat_compare_means(comparisons = my_comparisons) +
      scale_fill_brewer(name = NULL, palette = "Dark2",
                  label = c("1Q", "2Q", "3Q", "4Q"))

p_bp3 = df_alpha %>%
  filter(!is.na(BP3)) %>%
  ggplot(aes(x = BP3, y = shannon)) + 
  geom_boxplot(width = 0.5, aes(fill = BP3), alpha = 0.7) +
  geom_jitter(shape = 16, position = position_jitter(0.2), size = 0.7) + 
  labs(x = NULL, y = "Shannon diversity index", title = "BP3") + 
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "bottom") +
  stat_compare_means(comparisons = my_comparisons) +
      scale_fill_brewer(name = NULL, palette = "Dark2",
                  label = c("1Q", "2Q", "3Q", "4Q"))

p_alpha = ggarrange(p_mpb, p_ppb, p_bp3, 
          labels = c("A", "B", "C"),
          ncol = 2, nrow = 2, 
          common.legend = TRUE, legend = "bottom")
print(p_alpha)
ggsave(filename = "../figures/main figures/diversity/alpha.jpeg", plot = p_alpha,
       height = 8, width = 8, units = 'in', dpi = 300)
ggsave(filename = "../figures/main figures/diversity/alpha.pdf", plot = p_alpha,
       height = 8, width = 8)
```

## 2.2 Supplementary figures

```{r, fig.height=5, fig.width=6.25}
chem_list = c("TCS", "B_PB", "TCC", "DCP25", "EPB", "parabensum", 
              "DCP24", "BPA", "BPS", "BPF")
labels = c("TCS", "BPB", "TCC", "2,5-DCP", "EPB", "Combined Parabens", 
           "2,4-DCP", "BPA", "BPS", "BPF")
img_path = "../figures/supplementary figures/diversity/"

for (i in chem_list) {
  cat("\n \n \n")
  label = labels[which(chem_list == i)]
  
  df = df_alpha %>% filter(!is.na(!!as.name(i)))
  p = ggplot(data = df, aes(x = !!as.name(i), y = shannon)) + 
    geom_boxplot(width = 0.5, aes(fill = !!as.name(i)), alpha = 0.7) +
    geom_jitter(shape = 16, position = position_jitter(0.2), size = 0.7) + 
    labs(x = NULL, y = "Shannon diversity index", title = label) + 
    theme_bw() + 
    theme(plot.title = element_text(hjust = 0.5),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          legend.position = "bottom")
  
  n_level = nlevels(df[, i])
  if (n_level == 4) {
    my_comparisons = list(c("0", "1"), c("0", "2"), c("0", "3"),
                          c("1", "2"), c("1", "3"), c("2", "3"))
    p = p +
      stat_compare_means(comparisons = my_comparisons) +
      scale_fill_brewer(name = NULL, palette = "Dark2",
                  label = c("1Q", "2Q", "3Q", "4Q"))
  } else {
    p = p + stat_compare_means(label.x.npc = 0.5, label.y.npc = 1) +
      scale_fill_brewer(name = NULL, palette = "Set1")
  }
  print(p)
  cat("\n \n \n")
  ggsave(paste0(img_path, "alpha_", i, ".jpeg"), plot = p,
         height = 5, width = 6.25, units = 'in', dpi = 300)
  ggsave(paste0(img_path, "alpha_", i, ".pdf"), plot = p,
         height = 5, width = 6.25)
}
```

# 3. Beta diversity: Bray-Curtis dissimilarity

1. For the adult population.

2. P-values are obtained by Permutational Multivariate Analysis of Variance (PERMANOVA) and Permutational Analysis of Multivariate Dispersion (PERMDISP) for beta diversities.

## 3.1 Main figures

```{r, fig.height=7, fig.width=9}
chem_list = c("MPB", "PBP", "parabensum", "DCP24")
title_labels = c("MPB", "PPB", "Combined Parabens", "2,4-DCP")

m = matrix(c(1, 2, 3, 4, 5, 5), nrow = 3, ncol = 2, byrow = TRUE)
layout(mat = m, heights = c(0.45, 0.45, 0.1))
for (i in chem_list) {
  par(mar = c(1, 5, 4, 2))
  title_label = title_labels[which(chem_list == i)]
  panel_label = LETTERS[which(chem_list == i)]
  
  pseq = subset_samples(genus_data1, !is.na(get(i)))
  set.seed(123)
  # PERMANOVA
  permanova = adonis(t(abundances(pseq)) ~ get(i), 
                     data = meta(pseq), 
                     permutations = 999, method = "bray")$aov.tab
  
  # PERMDISP
  dis = vegdist(t(abundances(pseq)), method = "bray")
  groups = meta(pseq)[, i]
  n_level = nlevels(groups)
  
  groups = recode(groups, `0` = "1Q", `1` = "2Q", `2` = "3Q", `3` = "4Q")
  mod = betadisper(d = dis, group = groups, type = "median")
  p1 = signif(permanova$`Pr(>F)`[1], 2)
  p2 = signif(permutest(mod)$tab$`Pr(>F)`[1], 2)
  
  # Draw the Plot
  labs = paste0("PCoA", 1:2, " (", signif(100 * mod$eig / sum(mod$eig), 3), "%)")
  # brewer.pal(n = 8, name = "Dark2")
  plot(mod, pch = 15:18, cex.lab = 1.25, cex = 0.7, label.cex = 0.6, 
       main = title_label, xlab = labs[1], ylab = labs[2], 
       ylim = c(-0.5, 0.5), xlim = c(-0.6, 0.6), xaxt = "n",
       col = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A"), sub = NULL,
       hull = FALSE, ellipse = TRUE, conf = 0.68) # 68% data coverage for data ellipses
  axis(1, at = round(seq(-0.6, 0.6, by = 0.2), 1), las = 1)
  legend(x = 0.2, y = -0.3, cex = 0.7,
         legend = c(paste0("p (PERMANOVA) = ", p1),
                    paste0("p (PERMDISP) = ", p2)))
  mtext(text = panel_label, side = 3, line = 1.5, cex = 1.5, adj = -0.1)
}
par(mar = c(0, 0, 2, 0))
plot(1, type = "n", axes = FALSE, xlab = "", ylab = "")
legend(x = "top", inset = 0, lwd = 5, cex = .5, horiz = TRUE,
       legend = c("1Q", "2Q", "3Q", "4Q"), 
       col = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A"))

# JPEG
jpeg(filename = "../figures/main figures/diversity/beta.jpeg", 
     height = 7, width = 9, res = 300, units = "in")
m = matrix(c(1, 2, 3, 4, 5, 5), nrow = 3, ncol = 2, byrow = TRUE)
layout(mat = m, heights = c(0.45, 0.45, 0.1))
for (i in chem_list) {
  par(mar = c(1, 5, 4, 2))
  title_label = title_labels[which(chem_list == i)]
  panel_label = LETTERS[which(chem_list == i)]
  
  pseq = subset_samples(genus_data1, !is.na(get(i)))
  set.seed(123)
  # PERMANOVA
  permanova = adonis(t(abundances(pseq)) ~ get(i), 
                     data = meta(pseq), 
                     permutations = 999, method = "bray")$aov.tab
  
  # PERMDISP
  dis = vegdist(t(abundances(pseq)), method = "bray")
  groups = meta(pseq)[, i]
  n_level = nlevels(groups)
  
  groups = recode(groups, `0` = "1Q", `1` = "2Q", `2` = "3Q", `3` = "4Q")
  mod = betadisper(d = dis, group = groups, type = "median")
  p1 = signif(permanova$`Pr(>F)`[1], 2)
  p2 = signif(permutest(mod)$tab$`Pr(>F)`[1], 2)
  
  # Draw the Plot
  labs = paste0("PCoA", 1:2, " (", signif(100 * mod$eig / sum(mod$eig), 3), "%)")
  # brewer.pal(n = 8, name = "Dark2")
  plot(mod, pch = 15:18, cex.lab = 1.25, cex = 0.7, label.cex = 0.6, 
       main = title_label, xlab = labs[1], ylab = labs[2], 
       ylim = c(-0.5, 0.5), xlim = c(-0.6, 0.6), xaxt = "n",
       col = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A"), sub = NULL,
       hull = FALSE, ellipse = TRUE, conf = 0.68) # 68% data coverage for data ellipses
  axis(1, at = round(seq(-0.6, 0.6, by = 0.2), 1), las = 1)
  legend(x = 0.2, y = -0.3, cex = 0.7,
         legend = c(paste0("p (PERMANOVA) = ", p1),
                    paste0("p (PERMDISP) = ", p2)))
  mtext(text = panel_label, side = 3, line = 1.5, cex = 1.5, adj = -0.1)
}
par(mar = c(0, 0, 2, 0))
plot(1, type = "n", axes = FALSE, xlab = "", ylab = "")
legend(x = "top", inset = 0, lwd = 5, cex = .5, horiz = TRUE,
       legend = c("1Q", "2Q", "3Q", "4Q"), 
       col = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A"))
dev.off()

# PDF
pdf(file = "../figures/main figures/diversity/beta.pdf", height = 7, width = 9)
m = matrix(c(1, 2, 3, 4, 5, 5), nrow = 3, ncol = 2, byrow = TRUE)
layout(mat = m, heights = c(0.45, 0.45, 0.1))
for (i in chem_list) {
  par(mar = c(1, 5, 4, 2))
  title_label = title_labels[which(chem_list == i)]
  panel_label = LETTERS[which(chem_list == i)]
  
  pseq = subset_samples(genus_data1, !is.na(get(i)))
  set.seed(123)
  # PERMANOVA
  permanova = adonis(t(abundances(pseq)) ~ get(i), 
                     data = meta(pseq), 
                     permutations = 999, method = "bray")$aov.tab
  
  # PERMDISP
  dis = vegdist(t(abundances(pseq)), method = "bray")
  groups = meta(pseq)[, i]
  n_level = nlevels(groups)
  
  groups = recode(groups, `0` = "1Q", `1` = "2Q", `2` = "3Q", `3` = "4Q")
  mod = betadisper(d = dis, group = groups, type = "median")
  p1 = signif(permanova$`Pr(>F)`[1], 2)
  p2 = signif(permutest(mod)$tab$`Pr(>F)`[1], 2)
  
  # Draw the Plot
  labs = paste0("PCoA", 1:2, " (", signif(100 * mod$eig / sum(mod$eig), 3), "%)")
  # brewer.pal(n = 8, name = "Dark2")
  plot(mod, pch = 15:18, cex.lab = 1.25, cex = 0.7, label.cex = 0.6, 
       main = title_label, xlab = labs[1], ylab = labs[2], 
       ylim = c(-0.5, 0.5), xlim = c(-0.6, 0.6), xaxt = "n",
       col = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A"), sub = NULL,
       hull = FALSE, ellipse = TRUE, conf = 0.68) # 68% data coverage for data ellipses
  axis(1, at = round(seq(-0.6, 0.6, by = 0.2), 1), las = 1)
  legend(x = 0.2, y = -0.3, cex = 0.7,
         legend = c(paste0("p (PERMANOVA) = ", p1),
                    paste0("p (PERMDISP) = ", p2)))
  mtext(text = panel_label, side = 3, line = 1.5, cex = 1.5, adj = -0.1)
}
par(mar = c(0, 0, 2, 0))
plot(1, type = "n", axes = FALSE, xlab = "", ylab = "")
legend(x = "top", inset = 0, lwd = 5, cex = .5, horiz = TRUE,
       legend = c("1Q", "2Q", "3Q", "4Q"), 
       col = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A"))
dev.off() 
```

## 3.2 Supplementary figures

```{r, fig.height=5, fig.width=6.25}
chem_list = c("TCS", "B_PB", "TCC", "DCP25", "EPB", "BP3", "BPA", "BPF", "BPS")
labels = c("TCS", "BPB", "TCC", "2,5-DCP", "EPB", "BP3", "BPA", "BPF", "BPS")
img_path = "../figures/supplementary figures/diversity/"
for (i in chem_list) {
  cat("\n \n \n")
  label = labels[which(chem_list == i)]
  
  pseq = subset_samples(genus_data1, !is.na(get(i)))
  set.seed(123)
  # PERMANOVA
  permanova = adonis(t(abundances(pseq)) ~ get(i), 
                     data = meta(pseq), 
                     permutations = 999, method = "bray")$aov.tab
  
  # PERMDISP
  dis = vegdist(t(abundances(pseq)), method = "bray")
  groups = meta(pseq)[, i]
  n_level = nlevels(groups)
  
  # Iteration
  if (n_level == 4) {
    groups = recode(groups, `0` = "1Q", `1` = "2Q", `2` = "3Q", `3` = "4Q")
    mod = betadisper(d = dis, group = groups, type = "median")
    p1 = signif(permanova$`Pr(>F)`[1], 2)
    p2 = signif(permutest(mod)$tab$`Pr(>F)`[1], 2)
    
    # Draw the Plot
    labs = paste0("PCoA", 1:2, " (", signif(100 * mod$eig / sum(mod$eig), 3), "%)")
    # brewer.pal(n = 8, name = "Dark2")
    plot(mod, pch = 15:18, cex.lab = 1.25, cex = 0.7, label.cex = 0.6, 
         main = label, xlab = labs[1], ylab = labs[2], 
         ylim = c(-0.5, 0.5), xlim = c(-0.6, 0.6), xaxt = "n",
         col = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A"), sub = NULL,
         hull = FALSE, ellipse = TRUE, conf = 0.68) # 68% data coverage for data ellipses
    axis(1, at = round(seq(-0.6, 0.6, by = 0.2), 1), las = 1)
    legend(0.5, 0.4, legend = c("1Q", "2Q", "3Q", "4Q"),
           col = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A"), 
           pch = 15:16, cex = 0.8)
    legend(x = 0.2, y = -0.3, cex = 0.7,
           legend = c(paste0("p (PERMANOVA) = ", p1),
                      paste0("p (PERMDISP) = ", p2)))
    
    # Export
    jpeg(filename = paste0(img_path, "beta_", i, ".jpeg"), 
         height = 5, width = 6.25, res = 300, units = "in")
    plot(mod, pch = 15:18, cex.lab = 1.25, cex = 0.7, label.cex = 0.6, 
         main = label, xlab = labs[1], ylab = labs[2], 
         ylim = c(-0.5, 0.5), xlim = c(-0.6, 0.6), xaxt = "n",
         col = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A"), sub = NULL,
         hull = FALSE, ellipse = TRUE, conf = 0.68) # 68% data coverage for data ellipses
    axis(1, at = round(seq(-0.6, 0.6, by = 0.2), 1), las = 1)
    legend(0.5, 0.4, legend = c("1Q", "2Q", "3Q", "4Q"),
           col = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A"), 
           pch = 15:16, cex = 0.8)
    legend(x = 0.2, y = -0.3, cex = 0.7,
           legend = c(paste0("p (PERMANOVA) = ", p1),
                      paste0("p (PERMDISP) = ", p2)))
    dev.off()
    
    pdf(file = paste0(img_path, "beta_", i, ".pdf"), height = 5, width = 6.25)
    plot(mod, pch = 15:18, cex.lab = 1.25, cex = 0.7, label.cex = 0.6, 
         main = label, xlab = labs[1], ylab = labs[2], 
         ylim = c(-0.5, 0.5), xlim = c(-0.6, 0.6), xaxt = "n",
         col = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A"), sub = NULL,
         hull = FALSE, ellipse = TRUE, conf = 0.68) # 68% data coverage for data ellipses
    axis(1, at = round(seq(-0.6, 0.6, by = 0.2), 1), las = 1)
    legend(0.5, 0.4, legend = c("1Q", "2Q", "3Q", "4Q"),
           col = c("#1B9E77", "#D95F02", "#7570B3", "#E7298A"), 
           pch = 15:16, cex = 0.8)
    legend(x = 0.2, y = -0.3, cex = 0.7,
           legend = c(paste0("p (PERMANOVA) = ", p1),
                      paste0("p (PERMDISP) = ", p2)))
    dev.off()
  } else{
    mod = betadisper(d = dis, group = groups, type = "median")
    p1 = signif(permanova$`Pr(>F)`[1], 2)
    p2 = signif(permutest(mod)$tab$`Pr(>F)`[1], 2)
    
    # Draw the Plot
    labs = paste0("PCoA", 1:2, " (", signif(100 * mod$eig / sum(mod$eig), 3), "%)")
    # brewer.pal(n = 8, name = "Accent")
    plot(mod, pch = 15:16, cex.lab = 1.25, cex = 0.7, label.cex = 0.6, 
         main = label, xlab = labs[1], ylab = labs[2], 
         ylim = c(-0.5, 0.5), xlim = c(-0.6, 0.6), xaxt = "n",
         col = c("#E41A1C", "#377EB8"), sub = NULL,
         hull = FALSE, ellipse = TRUE, conf = 0.68) # 68% data coverage for data ellipses
    axis(1, at = round(seq(-0.6, 0.6, by = 0.2), 1), las = 1)
    legend(0.5, 0.4, legend = c("Below LOD", "Above LOD"),
           col = c("#E41A1C", "#377EB8"), 
           pch = 15:16, cex = 0.8)
    legend(x = 0.2, y = -0.3, cex = 0.7,
           legend = c(paste0("p (PERMANOVA) = ", p1),
                      paste0("p (PERMDISP) = ", p2)))
    
    # Export
    jpeg(filename = paste0(img_path, "beta_", i, ".jpeg"), 
         height = 5, width = 6.25, res = 300, units = "in")
    plot(mod, pch = 15:16, cex.lab = 1.25, cex = 0.7, label.cex = 0.6, 
         main = label, xlab = labs[1], ylab = labs[2], 
         ylim = c(-0.5, 0.5), xlim = c(-0.6, 0.6), xaxt = "n",
         col = c("#E41A1C", "#377EB8"), sub = NULL,
         hull = FALSE, ellipse = TRUE, conf = 0.68) # 68% data coverage for data ellipses
    axis(1, at = round(seq(-0.6, 0.6, by = 0.2), 1), las = 1)
    legend(0.5, 0.4, legend = c("Below LOD", "Above LOD"),
           col = c("#E41A1C", "#377EB8"), 
           pch = 15:16, cex = 0.8)
    legend(x = 0.2, y = -0.3, cex = 0.7,
           legend = c(paste0("p (PERMANOVA) = ", p1),
                      paste0("p (PERMDISP) = ", p2)))
    dev.off()
    
    pdf(file = paste0(img_path, "beta_", i, ".pdf"), height = 5, width = 6.25)
    plot(mod, pch = 15:16, cex.lab = 1.25, cex = 0.7, label.cex = 0.6, 
         main = label, xlab = labs[1], ylab = labs[2], 
         ylim = c(-0.5, 0.5), xlim = c(-0.6, 0.6), xaxt = "n",
         col = c("#E41A1C", "#377EB8"), sub = NULL,
         hull = FALSE, ellipse = TRUE, conf = 0.68) # 68% data coverage for data ellipses
    axis(1, at = round(seq(-0.6, 0.6, by = 0.2), 1), las = 1)
    legend(0.5, 0.4, legend = c("Below LOD", "Above LOD"),
           col = c("#E41A1C", "#377EB8"), 
           pch = 15:16, cex = 0.8)
    legend(x = 0.2, y = -0.3, cex = 0.7,
           legend = c(paste0("p (PERMANOVA) = ", p1),
                      paste0("p (PERMDISP) = ", p2)))
    dev.off()
  }
}
```

# Session information

```{r, message_group=FALSE, warning=FALSE, comment=NA}
sessionInfo()
```

